---
title: "Medical Incidents"
author: "Mahal Bilaoen - Edited by Nancy Carmona"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,dplyr,ggplot2,lubridate,brms, data.table, cmdstanr, season, dlnm, splines, survival, gnm)
 
```

04/27 11:00 AM - 11:10 AM - working on tmin analysis 

# Loading data 
```{r load.data}

# loading in data sets 

# load medical incidents from SFDPH
medical_incidents <- read.csv("Fire_Department_and_Emergency_Medical_Services_Dispatched_Calls_for_Service_20260309.csv")

# load all current temp data
daymet_current <- read.csv("daymet_current.csv")

# load 95th percentiles (zip code level) from historical data (30 year normal)
daymet_p95 <- read.csv("daymet_p95_byZIP.csv")



```


# Calculating AT 
```{r}

# first make a date column like in previous dataframe and add zero infront of site
daymet_current$site <- paste0("0", daymet_current$site)
daymet_p95$site <- as.character(daymet_p95$site)


# calculate AT from temp and vp 
daymet_current <- daymet_current %>%
  mutate(date = make_date(year = year) + days(yday - 1),
         month = month(date),
         apparent_tmin = -1.3 + 0.92*tmin + 2.2*(vp/1000),
         apparent_tmax = -1.3 + 0.92*tmax + 2.2*(vp/1000),
         apparent_tmean = (apparent_tmin+apparent_tmax)/2
         )


# merge data sets
daymet_current_join <- left_join(daymet_current, daymet_p95, by = "zip_code")

# remove duplicate observations
# Deduplicate by zip
daymet_heatwaves <- daymet_current_join %>%
  distinct(zip_code, date, .keep_all = TRUE) 


```

# Heatwave events 
```{r}

#tally heatwave days (above 95th percentile)
daymet_heatwaves <- daymet_heatwaves %>%
  filter(month >= 6 & month < 11) %>% #restrict to SF summer dates to avoid seasonal bias
  mutate(hw = apparent_tmean > p95_tmean) #trying mean now
daymet_heatwaves$zip_code <- as.character(daymet_heatwaves$zip_code)

daymet_heatwaves <- daymet_heatwaves %>%
  group_by(site.x, zip_code) %>%
  mutate(event_id = rleid(hw))

#summarise TRUE runs and 2+ day heatwaves
hw_events_2day <- daymet_heatwaves %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site.x, event_id, zip_code, year) %>%
  summarise(
    duration = n(),
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(apparent_tmean),
    .groups = "drop"
  ) %>%
  filter(duration >= 2) # enforce 2+ days

#summarise TRUE runs and 3+ day heatwaves
hw_events_3day <- daymet_heatwaves %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site.x, event_id, zip_code, year) %>%
  summarise(
    duration = n(), #adjust to just "duration" if 
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(apparent_tmean),
    .groups = "drop"
  ) %>%
  filter(duration >= 3) # enforce 3+ days

```





# Cleaning 911 calls 
```{r}
per_day <- medical_incidents |>
  mutate(Call.Date = mdy(Call.Date)) |>
  count(Zipcode.of.Incident, Call.Date) 

per_day <- per_day |>
  mutate(year = year(Call.Date),
         month = month(Call.Date),
         Zipcode.of.Incident = as.character(Zipcode.of.Incident)) |>
  relocate(year, .after = "Call.Date") |> 
  filter(month >= 6 & month < 11)

#seasonal bias to ONLY track assumed heatwave related days, need to find paper to support this 

#percentile_95 <- quantile(pertest$n, probs = 0.95, na.rm = TRUE)
#95th percentile is 44 calls for years 2018-2023
#95th percentile is 45 calls for years 2018-2023 BUT only through the summer months 

```




# Merging AT and 911 dfs
```{r}
overall_current <- left_join(
  per_day,
  daymet_heatwaves,
  by = c(
    "Call.Date" = "date",
    "Zipcode.of.Incident" = "zip_code",
    "year" = "year",
    "month" = "month"
  ),
  na_matches = "never"
)

overall_current <- overall_current %>% 
  mutate(day_abbr = format(as.Date(Call.Date), "%a"),
         date = Call.Date, 
        strata = interaction(month, day_abbr)) #create strata for cross basis

```

## NC: checking over dispersion

We used a conditional Poisson case-crossover model for daily 911 call counts. In these results, the coefficient for "hw" tells you how call counts differ on heat-wave days versus non-heat-wave days. The outcome of interest is a count of 911 models. The model uses a log link and the effects are interpreted as rate ratios after exponentiating coefficients. 

```{r}
library(dplyr)
library(lubridate)
library(gnm)
library(MASS)

dat <- overall_current %>%
  mutate(
    Call.Date = as.Date(Call.Date),
    year = lubridate::year(Call.Date),
    month = lubridate::month(Call.Date),
    dow = lubridate::wday(Call.Date, label = TRUE, abbr = TRUE),
    hw = as.integer(hw),
    cc_strata = interaction(Zipcode.of.Incident, year, month, dow, drop = TRUE)
  )

 fit_p <- gnm(
  n ~ hw,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat
)

pearson_dispersion <- sum(residuals(fit_p, type = "pearson")^2) / df.residual(fit_p)
deviance_dispersion <- deviance(fit_p) / df.residual(fit_p)

data.frame(
  pearson_dispersion = pearson_dispersion,
  deviance_dispersion = deviance_dispersion
)


beta <- coef(fit_p)["hw"]
se <- sqrt(vcov(fit_p)["hw", "hw"])

data.frame(
  term = "Heat wave",
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

# percent increase
(1.101841 - 1) * 100

# 10.18% higher 911 call rates on heat-wave days.


# 
c(
  lower_pct = (1.087843 - 1) * 100,
  upper_pct = (1.11602 - 1) * 100
)


```

# Results from Basic Model 

In conditional Poisson case-crossover models matched on ZIP code, year, month, and day of week, heat-wave days were associated with higher 911 call counts (IRR = 1.10, 95% CI: 1.09, 1.12). Overall, we find that heat-wave days were associated with a 10.2% increase in 911 calls (95% CI: 8.8% to 11.6%).


## Modeling only 1 day lag (no same day heat in model)

Heat-wave exposure on the previous day was also associated with elevated 911 call counts (IRR = 1.07, 95% CI: 1.05, 1.08), suggesting both immediate and short-term delayed effects.
 
```{r}

dat_lag1 <- dat %>%
  group_by(Zipcode.of.Incident) %>%
  arrange(Call.Date, .by_group = TRUE) %>%
  mutate(hw_lag1 = lag(hw, 1)) %>%
  ungroup()


fit_p_lag1 <- gnm(
  n ~ hw_lag1,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat_lag1
)

summary(fit_p_lag1)

beta <- coef(fit_p_lag1)["hw_lag1"]
se <- sqrt(vcov(fit_p_lag1)["hw_lag1", "hw_lag1"])

data.frame(
  term = "Heat wave (lag 1 day)",
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

```

## Model with same day heat + 1 day lag 

In conditional Poisson case-crossover models matched on ZIP code, year, month, and day of week, same-day heat-wave exposure was associated with higher 911 call counts (IRR = 1.09, 95% CI: 1.07, 1.11), and previous-day heat-wave exposure was also independently associated with an increase in 911 calls (IRR = 1.02, 95% CI: 1.01, 1.04). Most of the increase in 911 calls occurred on the heat-wave day itself, but there is still a small increase the following day.

```{r}


dat <- overall_current %>%
  mutate(
    Call.Date = as.Date(Call.Date),
    year = as.integer(format(Call.Date, "%Y")),
    month = as.integer(format(Call.Date, "%m")),
    dow = format(Call.Date, "%a"),
    hw = as.integer(hw),
    cc_strata = interaction(Zipcode.of.Incident, year, month, dow, drop = TRUE)
  ) %>%
  group_by(Zipcode.of.Incident) %>%
  arrange(Call.Date, .by_group = TRUE) %>%
  mutate(
    hw_lag1 = lag(hw, 1)
  ) %>%
  ungroup()

fit_lag01 <- gnm(
  n ~ hw + hw_lag1,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat
)

summary(fit_lag01)

beta <- coef(fit_lag01)[c("hw", "hw_lag1")]
se <- sqrt(diag(vcov(fit_lag01)))[c("hw", "hw_lag1")]

results <- data.frame(
  term = c("Heat wave (lag 0)", "Heat wave (lag 1)"),
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

results

```

#using tmin instead of tmean, creating a new df 
```{r}
#tally heatwave days (above 95th percentile)
tmin_heatwaves <- daymet_heatwaves %>%
  filter(month >= 6 & month < 11) %>% #restrict to SF summer dates to avoid seasonal bias
  mutate(hw = apparent_tmin > p95_tmin) #trying mean now
tmin_heatwaves$zip_code <- as.character(tmin_heatwaves$zip_code) 

tmin_heatwaves <- tmin_heatwaves %>%
  group_by(site.x, zip_code) %>%
  mutate(event_id = rleid(hw))

overall_tmin <- left_join(
  per_day,
  tmin_heatwaves,
  by = c(
    "Call.Date" = "date",
    "Zipcode.of.Incident" = "zip_code",
    "year" = "year",
    "month" = "month"
  ),
  na_matches = "never"
)

overall_tmin <- overall_tmin %>% 
  mutate(day_abbr = format(as.Date(Call.Date), "%a"),
         date = Call.Date, 
        strata = interaction(month, day_abbr)) #create strata for cross basis 
```

```{r}
dat_tmin <- overall_tmin %>%
  mutate(
    Call.Date = as.Date(Call.Date),
    year = lubridate::year(Call.Date),
    month = lubridate::month(Call.Date),
    dow = lubridate::wday(Call.Date, label = TRUE, abbr = TRUE),
    hw = as.integer(hw),
    cc_strata = interaction(Zipcode.of.Incident, year, month, dow, drop = TRUE)
  )

 fit_min <- gnm(
  n ~ hw,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat_tmin
)

pearson_min_dispersion <- sum(residuals(fit_p, type = "pearson")^2) / df.residual(fit_min)
deviance_min_dispersion <- deviance(fit_min) / df.residual(fit_min)

data.frame(
  pearson_dispersion = pearson_dispersion,
  deviance_dispersion = deviance_dispersion
)


beta <- coef(fit_p)["hw"]
se <- sqrt(vcov(fit_p)["hw", "hw"])

data.frame(
  term = "Heat wave",
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

# percent increase
(1.101841 - 1) * 100

# 10.18% higher 911 call rates on heat-wave days.


# 
c(
  lower_pct = (1.087843 - 1) * 100,
  upper_pct = (1.11602 - 1) * 100
)

```

