---
title: "Downloading NOAA Data"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  markdown: 
    wrap: 72
---

```{r set.knitr.options, echo=FALSE, message=FALSE, warning=FALSE}
#-----set knitr options-----

# clear environment
rm(list = ls())

# knitr options
suppressMessages(library(knitr))
opts_chunk$set(tidy = FALSE, echo = TRUE, message = FALSE)

# Load pacman, installing if needed and other packages as needed
if (!require("pacman")) { install.packages("pacman", repos = "http://cran.r-project.org") }

# Load other packages, installing as needed.
pacman::p_load(tidyverse, dplyr, httr, jsonlite, tigris, sf, gstat)


```

# About NOAA

The National Oceanic and Atmospheric Administration (NOAA) is a U.S.
federal agency that provides publicly available climate and weather
data, including long-term daily temperature, precipitation, and other
atmospheric variables. NOAA operates thousands of ground-based weather
stations across the country and maintains a large archive of historical
and near-real-time data.

## Using NOAA API

We will access NOAA data using the NOAA Climate Data Online (CDO) API,
which allows users to programmatically retrieve weather observations by
specifying parameters. To use the API, users must register for a free
API token from NOAA, and send HTTP requests with that token included.

In our project, we used the NOAA API to download daily temperature data
from two San Francisco stations for the period 2018–2015, which we will
then compared with Daymet estimates for validation.

This code will walk through obtaining weather data from NOAA API. First,
you will need to obtain an API Token to use the NOAA website.

```{r token, echo = FALSE}
# Set your NOAA API Token (replace with your own API key)
noaa_token <- ### 

```

### API Function

The function can be used based on the year, location id, and data type.
These inputs can be changed depending on what you would like to
download. The code below is using the FIPS code for San Francisco
county.

```{r api.function, eval = FALSE}

fetch_noaa_data_paginated <- function(start_date, end_date, location_id, datatypeids, token) {
  base_url <- "https://www.ncdc.noaa.gov/cdo-web/api/v2/data"
  all_results <- list()
  offset <- 1
  limit <- 1000
  more_data <- TRUE
  
  # Create the base query
  params <- list(
    datasetid = "GHCND",
    locationid = location_id,
    startdate = start_date,
    enddate = end_date,
    units = "metric",
    limit = limit,
    offset = offset
  )
  
  # Build the query string with multiple datatypeids
  datatypeid_params <- paste(paste0("datatypeid=", datatypeids), collapse = "&")
  
  while(more_data) {
    # Build the full URL with parameters
    full_url <- modify_url(
      base_url,
      query = c(params, list(offset = offset))
    )
    # Add the datatypeids manually
    full_url <- paste0(full_url, "&", datatypeid_params)
    
    response <- GET(full_url, add_headers(token = token))
    if (status_code(response) != 200) {
      cat("Error: Status", status_code(response), "\n")
      break
    }
    results <- fromJSON(content(response, "text", encoding = "UTF-8"))
    if (!is.null(results$results)) {
      all_results <- append(all_results, list(results$results))
      num_results <- length(results$results)
      cat("Fetched", num_results, "records (offset", offset, ")\n")
      offset <- offset + limit
      if (num_results < limit) {
        more_data <- FALSE
      }
    } else {
      more_data <- FALSE
    }
    Sys.sleep(1)
  }
  if (length(all_results) > 0) {
    return(bind_rows(all_results))
  } else {
    return(NULL)
  }
}


```

### Downloading data

Below we use the function created above to download data for the County
of San Francisco using it's FIPS code.

```{r use.fips.id, echo = FALSE}

# Example usage (loop by year, API works better if 1 year or less of data being requested at a time)
all_data <- list()
for (yr in 2000:2024) {
  cat("Downloading data for year:", yr, "\n")
  res <- fetch_noaa_data_paginated(
    start_date = paste0(yr, "-01-01"),
    end_date = paste0(yr, "-12-31"),
    location_id = "FIPS:06075", # FIPS for SF
    datatypeids = c("TMAX", "TMIN", "TAVG"), # Indicate the types of data being downloaded
    token = noaa_token
  )
  if (!is.null(res)) {
    all_data[[as.character(yr)]] <- res
  }
}

sf_weather <- bind_rows(all_data)


# save the data as a CSV to use for later 

write.csv(sf_weather, "sf_weather_noaa_2000_2024.csv", row.names = FALSE)

```

### Combine weather and station data

We then join weather data with station data, this will add the lat/long
data needed for next step in the interpolation.

```{r combine.data, echo = FALSE}

# read in downloaded csv of data 
# sf_weather <- read_csv("Climate Change/sf_weather_noaa_2000_2024.csv")


# get list of stations in san francisco 

response <- GET(
  url = "https://www.ncei.noaa.gov/cdo-web/api/v2/stations",
  query = list(
    locationid = "FIPS:06075",   # San Francisco city code
    datasetid = "GHCND",
    limit = 1000
  ),
  add_headers(token = noaa_token)
)

stations_json <- fromJSON(content(response, "text", encoding = "UTF-8"))
sf_stations_api <- stations_json$results

# Show the resulting stations
sf_station_ids <- sf_stations_api$id
sf_weather_only <- sf_weather %>% filter(station %in% sf_station_ids)
# Convert sf_stations_api to a tibble/data frame if not already
sf_stations_df <- as_tibble(sf_stations_api)

# Join weather data to station metadata (by station ID)
sf_weather_with_meta <- sf_weather_only %>%
  left_join(sf_stations_df, by = c("station" = "id"))


# Convert to Spatial Data
weather_points <- st_as_sf(sf_weather_with_meta, coords = c("longitude", "latitude"), crs = 4326)


# Get 2020 census tracts for San Francisco County (FIPS: 075)
sf_tracts <- tracts(state = "CA", county = "San Francisco", year = 2020, class = "sf")

# plot(sf_tracts$geometry)

sf_tracts <- st_transform(sf_tracts, crs = 4326)

