---
title: "Downloading Daymet Data"
output: html_document
date: "`r Sys.Date()`"
---

```{r set.knitr.options, include= FALSE, message=FALSE, warning=FALSE}
#-----set knitr options-----

# clear environment
rm(list = ls())

# knitr options
suppressMessages(library(knitr))
opts_chunk$set(tidy = FALSE, echo = TRUE, message = FALSE)

# Load pacman, installing if needed and other packages as needed
if (!require("pacman")) { install.packages("pacman", repos = "http://cran.r-project.org") }

# Load other packages, installing as needed.
pacman::p_load(tidyverse, dplyr, httr, jsonlite, tigris, sf, gstat, daymetr)

```

# About Daymet

Unlike NOAA (National Oceanic and Atmospheric Administration), which uses observational climate and weather data collected from ground stations, satellites, and other sources, Daymet provides gridded daily surface weather data (e.g., temperature, precipitation) for North America, derived from interpolation of ground-based station data, and is especially useful for ecological and environmental modeling where spatially continuous data are needed.

## Using Daymet R package

We will use the *daymetr* package to download weather data from 2018-2025. More information and examples of how to use the R package can be found here: <https://cran.r-project.org/web/packages/daymetr/vignettes/daymetr-vignette.html>

To download the data we need to select the locations of the estimates. Given we want to examine weather at the census tract level, we can use census tract centroids to extract estimates at these locations.

A census tract centroid is the geographic center point (latitude and longitude) of a census tract, which is a small, relatively permanent subdivision of a county used for statistical purposes. The centroid represents the approximate "middle" of the tract and is often used to link spatial data (like climate or health data) to the tract's location.

First, we need to create a CSV file with information needed for batch downloading using the *daymetr* package.

```{r centroids, echo = FALSE}

# To use Daymet we will need to make csv file with : site (GEOID), latitude, longitude

# Get 2020 census tracts for San Francisco County (FIPS: 075)
sf_tracts <- tracts(state = "CA", county = "San Francisco", year = 2020, class = "sf")

# extract centroids 
tract_centroids <- st_centroid(sf_tracts)

# select the variables 
sf_centroids <- tract_centroids %>% select(GEOID, INTPTLAT, INTPTLON)

# remove geometry 
sf_centroid_df <- st_drop_geometry(sf_centroids)

# change to numeric 
sf_centroid_df <- sf_centroid_df %>% mutate(across(c(INTPTLAT, INTPTLON), as.numeric))

# rename to use with Daymet function 
sf_centroid_df <- sf_centroid_df %>% 
                        rename(site = GEOID,
                               lat = INTPTLAT,
                               lon = INTPTLON)


# save a csv with the centroid locations for San Francisco 
# write.csv(sf_centroid_df, "sf_centroids.csv", row.names = FALSE)


```

### Batch downloads

The download_daymet_batch function from the daymetr R package allows users to download Daymet data for multiple locations in a single batch. Users provide a table of site information (with site names, latitude, and longitude), and the function retrieves daily weather data (e.g., temperature, precipitation) for each site over specified years. It’s useful for efficiently collecting climate data for many locations at once.

```{r daymet.download, eval = TRUE}

# warning: downloading data takes a LONG time to run, so only run once, set chunk to eval = FALSE to help prevent it from running again 

# download batch data using SF centroid locations
# make sure to indicate the year, end
# the argument "simplify" can be used to create a dataframe of the data, not using it will result your data to be in a list format 

df_batch <- download_daymet_batch(file_location = "sf_centroids.csv",
start = 2000,
end = 2024,
simplify = TRUE)


```

### Data wrangling

We now need to fix the names of the measurement types.

```{r data.wrangling, echo = FALSE}

# what are the names of each category? use unique to find the names  
unique(sf_daymet$measurement)

sf_daymet <- sf_daymet %>%
  mutate(measurement = recode(measurement,
    "tmax..deg.c." = "tmax",
    "tmin..deg.c." = "tmin",
    "prcp..mm.day." = "prcp",
    "srad..W.m.2." = "srad",
    "swe..kg.m.2." = "swe",
    "vp..Pa." = "vp",
    "dayl..s." = "dayl"
  ))

# check names again 
unique(sf_daymet$measurement)

# check dimensions, quick check of the number of rows 
dim(sf_daymet) # 14839442

# remove the level that does not have estimates 
sf_daymet <- sf_daymet %>%
  filter(!grepl("outside DAYMET spatial coverage", measurement))

# check dimensions, quick check of the number of rows 
dim(sf_daymet) # 14839440

# check names again 
unique(sf_daymet$measurement)

# save a csv with the cleaned up data 
# write.csv(sf_daymet, "Data/sf_daymet_clean.csv", row.names = FALSE)

```

