---
title: "mapping"
author: "Mahal Bilaoen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,dplyr,ggplot2,sf,tmap, tibble, patchwork,tigris, data.table)
```

#reading in data 
```{r}
#all svi data 
svi <- read.csv("SVI_2022_Census_Tract.csv")

#2018-23 temperature data
daymet_current <- read.csv("daymet_current.csv")

#p95 by census tract, done separately from zip code level data for accuracy
daymet_p95 <- read.csv("daymet_p95_byCENSUS.csv")
```

#reading in shapefile
```{r}
shapefile_path <- "C:/Users/meme1/OneDrive/Documents/PH 699 Bilaoen/Census Tract shp"

shape_data <- st_read(shapefile_path)

shape_data <- shape_data %>% 
  filter(aland > 0, #land area only
         st_coordinates(st_centroid(geometry))[,1] > -123) #removes weird water tracts, not perfect

#head(shape_data)

#plot(shape_data["geometry"])
```

#filtering for SF svi 
```{r}
#filtering through SVI data for san francisco county only, renaming FIPS to site to match centroids, removing other CA related labels 
svi <- svi %>% 
  filter(COUNTY == "San Francisco County") %>% 
  rename(site = FIPS) %>% 
  dplyr::select(where(~ !is.character(.))) %>% 
  dplyr::select(-ST, -STCNTY) %>% 
  filter(site != "6075980401")

#according to SVI documentation, a number of -999 means that there was no data available or the data was NULL 
svi <- svi %>% 
  mutate(
    RPL_THEMES = na_if(RPL_THEMES,-999), 
    RPL_THEME1 = na_if(RPL_THEME1, -999), 
    RPL_THEME2 = na_if(RPL_THEME2, -999), 
    RPL_THEME3 = na_if(RPL_THEME3, -999), 
    RPL_THEME4 = na_if(RPL_THEME4, -999)) %>%
  dplyr::select(site,RPL_THEMES,RPL_THEME1,RPL_THEME2,RPL_THEME3,RPL_THEME4) 
```

#Calculating AT
```{r}
daymet_p95$site <- as.character(daymet_p95$site)
daymet_current$site <- as.character(daymet_current$site)


# calculate AT from temp and vp 
daymet_current <- daymet_current %>%
  mutate(date = make_date(year = year) + days(yday - 1),
         month = month(date),
         apparent_tmin = -1.3 + 0.92*tmin + 2.2*(vp/1000),
         apparent_tmax = -1.3 + 0.92*tmax + 2.2*(vp/1000),
         apparent_tmean = (apparent_tmin+apparent_tmax)/2
         )

# merge data sets
daymet_current_join <- left_join(daymet_current, daymet_p95, by = "site")

# remove duplicate observations
# Deduplicate by census tracs
daymet_heatwaves <- daymet_current_join %>%
  distinct(site, date, .keep_all = TRUE) 
```

#calculating 2+ day events 
- using only 2 day events for base idea/concept, can be changed to 3 day if needed 
```{r}

#tally heatwave days (above 95th percentile)
daymet_heatwaves <- daymet_heatwaves %>%
  mutate(hw = apparent_tmin > p95_min) #using at-min for people level data
daymet_heatwaves$zip_code <- as.character(daymet_heatwaves$zip_code)

daymet_heatwaves <- daymet_heatwaves %>%
  group_by(site) %>%
  mutate(event_id = rleid(hw))

#summarise TRUE runs and 2+ day heatwaves
hw_events_2day <- daymet_heatwaves %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, year) %>%
  summarise(
    duration = n(),
    start_date = min(date),
    end_date = max(date),
    mean_tmin = mean(apparent_tmin),
    .groups = "drop"
  ) %>%
  filter(duration >= 2) # enforce 2+ days
```

#yrly 2 day hw events
```{r}
hw_yearly_2day <- hw_events_2day %>%
  mutate(year = lubridate::year(start_date)) %>%
  group_by(site, year) %>%
  summarise(
    #n_events = n(),
    total_hw_days = sum(duration),
    #total_hw_events = sum(n_events),
    yrly_meantmin = mean(mean_tmin),
    .groups = "drop"
  )

hw_2day_occurences <- hw_yearly_2day %>% 
  group_by(site) %>% 
  summarize(
    mean_hw_2day = mean(total_hw_days))
```

#joining heatwave and svi data 
```{r}
svi$site <- as.character(svi$site)

svi_heat <- left_join(svi,hw_2day_occurences, by = "site") 

svi_heat <- svi_heat %>% 
  ungroup() %>% 
  mutate(
    site = paste0("0", site),
    svi_quartile = ntile(RPL_THEMES, 4),     # Q1 = least vulnerable, Q4 = most 
    heat_quartile = ntile(mean_hw_2day, 4))      # Q1 = least vulnerable, Q5 = most 
```

#joining svi+hw df to census tract shp, creating bivar column
-- from here mostly using code from PINC 2025 to create bivar map 
```{r}
#joining hw+svi df to census shp
svi_shp <- left_join(svi_heat,shape_data, by = c("site" = "geoid"))

#renaming values/column names to read in all data easier 
tracts_bivar_all <- svi_shp %>% 
  pivot_longer( 
    cols = starts_with("mean_hw_"),  
    names_to = "definition", 
    values_to = "heat_value" 
  ) 

#create a bivariate class between svi and heat
tracts_bivar_all <- tracts_bivar_all %>% 
  mutate( 
    bivar_class = paste0(svi_quartile, "-", heat_quartile) 
  )

#
tracts_bivar_all <- tracts_bivar_all %>% 
  mutate(definition = recode(definition, 
    "mean_hw_2day"   = "2-Day ≥ 95th"
  )) %>%  
  filter(definition == "2-Day ≥ 95th") #only looking at one heatwave definition 
```

```{r}

#forcing r to read in the df as a shape/mappable object 
tracts_bivar_all <- st_as_sf(tracts_bivar_all)

# Use the same bivar_palette you used in the map 

bivar_palette <- c( 
  "1-1" = "#f2f0f7", "1-2" = "#cbc9e2", "1-3" = "#9e9ac8", "1-4" = "#756bb1", 
  "2-1" = "#e7d4e8", "2-2" = "#d1b9d8", "2-3" = "#b39dd0", "2-4" = "#8e6cbf", 
  "3-1" = "#fbb4b9", "3-2" = "#f768a1", "3-3" = "#dd3497", "3-4" = "#ae017e", 
  "4-1" = "#fcae91", "4-2" = "#fb6a4a", "4-3" = "#ef3b2c", "4-4" = "#99000d" 
) 

# Create 4x4 grid 

bivar_legend_df <- expand.grid( 
  svi_q = 1:4, 
  heat_q = 1:4 
) %>% 
  mutate( 
    bivar_class = paste0(svi_q, "-", heat_q), 
    fill = bivar_class  # match to your map's fill values 
  ) 


bivar_legend <- ggplot(bivar_legend_df, aes(x = heat_q, y = svi_q, fill = bivar_class)) + 
  geom_tile() + 
  scale_fill_manual(values = bivar_palette, guide = "none") + 
  scale_x_continuous( 
    breaks = 1:4, 
    labels = c("Low", "", "", "High"), 
    expand = c(0, 0) 
  ) + 
  scale_y_continuous( 
    breaks = 1:4, 
    labels = c("Low", "", "", "High"), 
    expand = c(0, 0) 
  ) + 
  labs(x = "Heatwave Exposure", y = "Social Vulnerability") + 
  coord_fixed() + 
  theme_minimal() + 
  theme( 
    axis.title = element_text(size = 8, face = "bold"), 
    axis.text = element_text(size = 8), 
    panel.grid = element_blank(), 
    plot.margin = margin(2, 2, 2, 2) 
  ) 

 
main_map <- ggplot(tracts_bivar_all) + 
  geom_sf(aes(fill = bivar_class), color = "white", size = 0.2) + 
  scale_fill_manual(values = bivar_palette, guide = "none") + # legend removed 

# if you want to facet wrap facet_wrap(~ definition, ncol = 2) + 

  labs( 
    title = "Bivariate Map of Social Vulnerability and Heat Exposure from 2018-2023", 
    subtitle = "Quartile Classification of SVI and Heatwave Metrics, Census Tract Level",   
    #caption = "Figure 2" 
    ) + 
  theme_void() + 
  theme( 
    plot.title = element_text(face = "bold", hjust = 0.5), 
    strip.text = element_text(face = "bold", size = 12, hjust = 0.5), 
    plot.caption = element_text(size = 8, face = "italic", hjust = 0.5, margin = margin(b=10)), plot.margin = margin(r=10) 
  ) 

final_plot <- main_map +  
  inset_element(bivar_legend, left = -0.5, bottom = 0.02, right = 0.25, top = 0.25) 


print(final_plot) 
```

#saving a picture of the map as a png 
```{r}
ggsave("Svi_Heat_2018_2023_Census.png", plot = final_plot, width = 9, height = 6, dpi = 300)
```

