---
title: "Medical Incidents"
author: "Mahal Bilaoen"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,dplyr,ggplot2,lubridate,brms, data.table, cmdstanr, season)

#if you want to use cmdstanr to run code without having to use C++, is supposed to help run MCMC chains faster 
# install.packages("remotes")
# remotes::install_github("stan-dev/cmdstanr")

#season package is specifically for case cross design 
```

## Bayesian modeling

In an attempt to understand which neighborhoods are potentially the most vulnerable due to heatwaves, I will be utilizing a bayesian linear regression model to use medical incidents as a proxy for vulnerability. The goal is to understand where there are the least air conditioning resources are available to people based on medical incidents, assuming that people who are calling in are in dire need of medical care due to heat. 

## Hypothesis: There are more medical incident 911 calls on heatwave days than on non heatwave days. 

independent - heat / temperature 
dependent - phone calls / medical incidents 
confounding (?) - indoor heat / location of the city 

It should be noted that this is only looking at calls on heatwave days, and clearer data on the type of medical incident days after a heatwave may be more inclusive of heatwave related injuries/exposure. Given that the public data is vague, we can only make inferences about the types of 911 medical incident calls being made. Other literature states that they included data that were not directly related to heat, as heat can exacerbate present illness/issues that people may be experiencing. If there was clearer available on the reasoning for the call, then there wouldn't be as deep of a model utilized here. 

#methods
In order to determine if the hypothesis is true, I will be utilizing the brm function from the "brms" package to determine a baseline of medical incidents. Using a bayesian model will help determine the likelihood/probability of medical incidents happening during a heatwave vs not a heatwave to get a baseline idea of how many 911 calls are happening. Using a 5 year period to test this information was used due to limited computational abilities. If a 20 year analysis could be completed that would the most ideal. 

#loading in data 
```{r}
#loading in data sets 
#medical incidents 
medical_incidents <- read.csv("Fire_Department_and_Emergency_Medical_Services_Dispatched_Calls_for_Service_20260309.csv")

daymet_current <- read.csv("daymet_current.csv")

```

04/05 
6:20 PM - working on case crossover design 

#cleaning 911 calls 
```{r}
per_day <- medical_incidents %>%
  mutate(Call.Date = mdy(Call.Date)) %>%
  count(Zipcode.of.Incident, Call.Date)

per_day <- per_day %>% 
  mutate(Year = year(Call.Date)) %>%
  relocate(Year, .after = "Call.Date")

pertest <- per_day
#duplicate data frame in case something i code is wrong 

pertest <- pertest %>% 
  mutate(month = month(Call.Date)) %>% 
  filter(month >= 6 & month < 11) 
#seasonal bias to ONLY track assumed heatwave related days, need to find paper to support this 

percentile_95 <- quantile(pertest$n, probs = 0.95, na.rm = TRUE)
#95th percentile is 44 calls for years 2018-2023
#95th percentile is 45 calls for years 2018-2023 BUT only through the summer months 

 

```

#calculating apparent temp 
```{r}
daymet_current$site <- paste0("0", daymet_current$site)

daymet_current <- daymet_current %>%
  mutate(date = make_date(year = year) + days(yday - 1),
         month = month(date),
         p95 = 18.7, #95th percentile for AT 
         tmean = (tmin+tmax)/2, 
         vp = vp/1000) 

daymet_current <- daymet_current %>%
  mutate(AT = (-1.3 + 0.92*(tmean) + 2.2*(vp)))

daymet_heatwaves <- daymet_current %>%
  distinct(site, date, .keep_all = TRUE)

daymet_heatwaves <- daymet_heatwaves %>%
  filter(month >= 6 & month < 11) %>% #restrict to SF summer dates to avoid seasonal bias
  mutate(hw = tmean > p95)

daymet_heatwaves <- daymet_heatwaves %>%
  #arrange(site, date) %>%
  group_by(site) %>%
  mutate(event_id = rleid(hw)) 

#summarise TRUE runs and 2+ day heatwaves
hw_events_2day <- daymet_heatwaves %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(),
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(tmean),
    AT = AT , 
    .groups = "drop"
  ) %>%
  filter(duration >= 2) # enforce 2+ days

#summarise TRUE runs and 3+ day heatwaves
hw_events_3day <- daymet_heatwaves %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(), #adjust to just "duration" if 
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(tmean),
    AT = AT , 
    .groups = "drop"
  ) %>%
  filter(duration >= 3) # enforce 3+ days
```

#yearly AT outcomes 
Would it be possible to run this model on a yearly basis (counts of hw days vs non hw days)? In order to avoid running extremely long models?  
```{r}
hw_yearly_3day <- hw_events_3day %>%
  mutate(year = lubridate::year(start_date)) %>%
  group_by(site, year, zip_code) %>%
  summarise(
    #n_events = n(),
    total_hw_days = sum(duration),
    #total_hw_events = sum(n_events),
    #avg_duration = mean(duration),
    #max_duration = max(duration),
    yrly_meantemp = mean(mean_tmean),
    .groups = "drop"
  )

hw_yearly_2day <- hw_events_2day %>%
  mutate(year = lubridate::year(start_date)) %>%
  group_by(site, year, zip_code) %>%
  summarise(
    #n_events = n(),
    total_hw_days = sum(duration),
    #total_hw_events = sum(n_events),
    #avg_duration = mean(duration),
    #max_duration = max(duration),
    yrly_meantemp = mean(mean_tmean),
    .groups = "drop"
  )
```

#counting the amount of 911 calls 
-should expect numbers to be within this range of the model, no actual idea on sd   
```{r}
summary(test)
# n min = 1 , max = 95, med = 8, mean = 12.46 

hist(pertest$n) 
#distribution is mostly from 0 to 20 calls per day over a 5 year average 

```

#merging AT and 911 dfs
```{r}
overall <- left_join(
  pertest,
  daymet_heatwaves,
  by = c("Call.Date" = "date", "Zipcode.of.Incident" = "zip_code"),
  na_matches = "never"
)

```

#bayesian model for at and 911 calls 
need to also consider lag 

For more information about the brms package: 
https://mspeekenbrink.github.io/sdam-r-companion/bayesian-estimation-with-brms.html 
```{r}

library(cmdstanr)
#install_cmdstan()
set_cmdstan_path(path = NULL)

fit <- brm(
  n ~ hw + (1 | Zipcode.of.Incident), #the outcome ~ the exposure, grouping by zipcode 
  data = overall, 
  family = negbinomial(), #negbinomial is for true/false counts   
   prior = c( #priors were set after running test runs of the model 
    prior(normal(0,1), class = "b"), #percentage affect of hw 
    prior(exponential(1), class = "sd"), #standard deviation
    prior(exponential(1), class = "shape"), #spatial distribution 
    prior(normal(2,1), class = "Intercept")), #what is your baseline calls on a false day     
  chains = 4, #how many times running MC 
  cores = 4, #how fast ur machine can run the model CPU 
  threads = threading(2), #forces iterations to be run at the same time much faster, MIGHT mess with coefficients  
  iter = 2000, #how many samples are you taking 
  warmup = 1000, #"throwaway" sample size 
  backend = "cmdstanr", 
  control = list(adapt_delta = 0.95),
#  sample_prior = TRUE , #sampling prior while also fitting the model 
#  seed = 2149 #reproducible sampling #godspeed  
  ) 

#saving the model after it's been run for future analysis
#saveRDS(fit, "post_bayesian.rds")

# bayesian <- readRDS("911_bayesian.rds") 

```
04/01 - this is not a funny joke ;_;
 Family: negbinomial 
  Links: mu = log 
Formula: n ~ hw + (1 | Zipcode.of.Incident) 
   Data: overall (Number of observations: 208792) 
  Draws: 4 chains, each with iter = 2000; warmup = 1000; thin = 1;
         total post-warmup draws = 4000

Multilevel Hyperparameters:
~Zipcode.of.Incident (Number of levels: 25) 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)     0.96      0.15     0.73     1.31 1.03      113      145

Regression Coefficients:
          Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept     2.11      0.22     1.72     2.55 1.20       14      118 ->rhat still too high 
hwTRUE        0.07      0.00     0.06     0.07 1.00     2346     1739

Further Distributional Parameters:
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
shape    57.51      0.71    56.18    58.93 1.00      912     1492

03/25
Family: negbinomial 
  Links: mu = log 
Formula: n ~ hw + (1 | Zipcode.of.Incident) 
   Data: overall (Number of observations: 208792) 
  Draws: 4 chains, each with iter = 500; warmup = 250; thin = 1;
         total post-warmup draws = 1000

Multilevel Hyperparameters:
~Zipcode.of.Incident (Number of levels: 25) 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)     0.93      0.13     0.70     1.23 1.15       23       89

Regression Coefficients:
          Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept     2.06      0.19     1.75     2.49 1.80        6       13  <- rhat >1
hwTRUE        0.07      0.00     0.06     0.07 1.01      782      591

Further Distributional Parameters:
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
shape    57.52      0.74    56.10    59.07 1.01      397      381



#bayesian model data 
```{r}
bayesian <- readRDS("911_bayesian.rds") 
summary(fit)
plot(fit) #i am not entirely sure how to read these graphs 
 

```

# Case Crossover Design 
Due to a lack of ability to obtain significant coefficients in (rhat >1), switching over to a casecross design will give insights into whether or not medical incident type calls increases when a heatwave occurs. This does not paint a complete picture in defining a baseline amount of 911 calls as a Bayesian model would. However, this type of study design is useful for looking at overall trends. 

To test this hypothesis out, we will begin with the year 2017 as it was one of the hottest heatwaves on record in San Francisco. 

04/06 
10:25 am - working on getting 2017 heatwave data with AT, using tmean 

#medical calls
```{r}
only_2017 <- read.csv("2017_data.csv")

only_2017 <- only_2017 %>%
  mutate(Call.Date = mdy(Call.Date)) %>%
  count(Zipcode.of.Incident, Call.Date)

only_2017 <- only_2017 %>% 
  mutate(Year = year(Call.Date)) %>%
  relocate(Year, .after = "Call.Date")

# testonly_2017 <- only_2017
#duplicate data frame in case something i code is wrong 

# testonly_2017 <- testonly_2017 %>%   , useful for yr by yr comparisons like the houston paper  
#  mutate(month = month(Call.Date)) %>% 
#  filter(month >= 6 & month < 11)  

```
 
#heatwaves 
```{r}
baseline <- read.csv("daymet_baseline.csv")

# baseline$site <- paste0("0", baseline$site) , saving incase we get census level data

baseline <- baseline %>%
  mutate(date = make_date(year = year) + days(yday - 1),
         month = month(date),
         p95 = 18.7,#95th percentile for AT for baseline yrs 
         tmean = (tmin+tmax)/2,
         vp = vp/1000,
         AT = (-1.3 + 0.92*(tmean) + 2.2*(vp)),
         hw95 = tmean >= p95 #starting with mean first then work on tmin analysis?
         )

baseline <- baseline %>%
  distinct(zip_code, date, .keep_all = TRUE) %>% #keeping by zipcode since 911 data is by zipcode
  relocate(date, .after = yday)

hw_2017 <- baseline %>%
  filter(year == 2017)
  

#hw_2017 <- hw_2017 %>%
#  filter(month >= 6 & month < 11) %>% #restrict to SF summer dates to avoid seasonal bias
#  mutate(hw = tmin > p95) 
```

```{r}
overall <- left_join(
  only_2017,
  hw_2017,
  by = c("Call.Date" = "date", "Zipcode.of.Incident" = "zip_code"),
  na_matches = "never"
) 

overall <- overall %>% 
  mutate(day_abbr = format(as.Date(Call.Date), "%a"),
         date = Call.Date) %>% #adding day of week to match cases/controls 
  relocate(tmean, hw95,day_abbr, .after = n) #visual sanity

#mon tue wed thu fri sat sun 
```

#casecross model
incorporate a 3 day lag like the cdph --> states that heat related stroke deaths can occur days later than the initial heatwave 

11:13 am date not in date format? 

12:36 pm 1:48 pm - do 2018 - 2023 data ? 
```{r}
#day of week case cross 
model2 = casecross(n ~ hw95+tmean, matchdow = TRUE, data = overall)
summary(model2)
head(model2)
```
Time-stratified case-crossover with a stratum length of 28 days
Matched on day of the week
Total number of cases 113939 
Number of case days with available control days 364 
Average number of control days per case day 3 

Parameter Estimates:
                 coef exp(coef)    se(coef)          z     Pr(>|z|)
hw95TRUE -0.014804682 0.9853044 0.017735190 -0.8347631 4.038511e-01 <- not sig BUT only within a yr
tmean     0.008534511 1.0085710 0.002188718  3.8993190 9.646363e-05 
