---
title: "Sensitivity"
output: html_document
---

The purpose of this document is to look at the percentage differences in IRR of medical incidents (911 calls) when Apparent Temperature is set at different baselines (85th, 97.5th). In the original Case_crossover code, 95th mean and minimum historical baseline temperatures were taken. 

#setting up packages 
```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,dplyr,ggplot2,lubridate,brms, data.table, cmdstanr, season, dlnm, splines, survival, gnm, broom, MASS)
```

#loading in data 
```{r load.data}
# load medical incidents from SFDPH
medical_incidents <- read.csv("~/PH 699 Bilaoen/Fire_Department_and_Emergency_Medical_Services_Dispatched_Calls_for_Service_20260309.csv")

# load all current temp data
daymet_current <- read.csv("~/PH 699 Bilaoen/daymet_current.csv")

# load percentile (zip code level) from historical data (30 year normal) - includes min/mean/max of 85th/95/97.5th 
#limited to months 4-10 
daymet_percentiles <- read.csv("~/PH 699 Bilaoen/NOVEMBER/daymet_percentile_byZIP_stratified.csv") 
```

#cleaning data 
Since there are multiple data sets being created, several of the original verbiage in the case crossover code has been shortened for ease of reading for this code: 

- hw = heatwave, calculated by the current temperature data measured against historical baseline with         respect to the percentile
- hwe = heatwave event, shortened due to long data frame names, calculated looking at how long a             heatwave lasts (2+ or 3+)

##calculating AT 
```{r}
# calculate AT from temp and vp 
daymet_current <- daymet_current %>%
  mutate(date = make_date(year = year) + days(yday - 1),
         month = month(date),
         apparent_tmin = -1.3 + 0.92*tmin + 2.2*(vp/1000),
         apparent_tmax = -1.3 + 0.92*tmax + 2.2*(vp/1000),
         apparent_tmean = (apparent_tmin+apparent_tmax)/2
         )

# merge data sets
daymet_current_join <- left_join(daymet_current, daymet_percentiles, by = "zip_code")

# remove duplicate observations
# Deduplicate by zip
daymet_heatwaves <- daymet_current_join %>%
  distinct(zip_code, date, .keep_all = TRUE) 
```

##85th mean heatwave days + events 
```{r}
#tally heatwave days at 85th percentile mean
hw_85_mean <- daymet_heatwaves %>%
  filter(month >= 4 & month < 10) %>% #restrict to SF summer dates to avoid seasonal bias
  mutate(hw = apparent_tmean > p85_mean) 
hw_85_mean$zip_code <- as.character(hw_85_mean$zip_code)

hw_85_mean <- hw_85_mean %>%
  group_by(site, zip_code) %>%
  mutate(event_id = rleid(hw))

#summarise TRUE runs and 2+ day heatwaves
hwe_2day_85_mean <- hw_85_mean %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(),
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(apparent_tmean),
    .groups = "drop"
  ) %>%
  filter(duration >= 2) # enforce 2+ days

#summarise TRUE runs and 3+ day heatwaves
hwe_3day_85_mean <- hw_85_mean %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(), #adjust to just "duration" if 
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(apparent_tmean),
    .groups = "drop"
  ) %>%
  filter(duration >= 3) # enforce 3+ days
```

###85th minimum heatwave days + events 
```{r}
#tally heatwave days at 85th percentile mean
hw_85_min <- daymet_heatwaves %>%
  filter(month >= 4 & month < 10) %>% #restrict to SF summer dates to avoid seasonal bias
  mutate(hw = apparent_tmin > p85_min) 

hw_85_min$zip_code <- as.character(hw_85_min$zip_code)

hw_85_min <- hw_85_min %>%
  group_by(site, zip_code) %>%
  mutate(event_id = rleid(hw))

#summarise TRUE runs and 2+ day heatwaves
hwe_2day_85_min <- hw_85_min %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(),
    start_date = min(date),
    end_date = max(date),
    mean_tmin = mean(apparent_tmin),
    .groups = "drop"
  ) %>%
  filter(duration >= 2) # enforce 2+ days

#summarise TRUE runs and 3+ day heatwaves
hwe_3day_85_mean <- hw_85_mean %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(), #adjust to just "duration" if 
    start_date = min(date),
    end_date = max(date),
    mean_tmin = mean(apparent_tmin),
    .groups = "drop"
  ) %>%
  filter(duration >= 3) # enforce 3+ days
```

##97.5th mean heatwave days + events 
```{r}
#tally heatwave days at 85th percentile mean
hw_975_mean <- daymet_heatwaves %>%
  filter(month >= 4 & month < 10) %>% #restrict to SF summer dates to avoid seasonal bias
  mutate(hw = apparent_tmean > p97.5_mean) 

hw_975_mean$zip_code <- as.character(hw_975_mean$zip_code)

hw_975_mean <- hw_975_mean %>%
  group_by(site, zip_code) %>%
  mutate(event_id = rleid(hw))

#summarise TRUE runs and 2+ day heatwaves
hwe_2day_975_mean <- hw_975_mean %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(),
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(apparent_tmean),
    .groups = "drop"
  ) %>%
  filter(duration >= 2) # enforce 2+ days

#summarise TRUE runs and 3+ day heatwaves
hwe_3day_975_mean <- hw_975_mean %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(), #adjust to just "duration" if 
    start_date = min(date),
    end_date = max(date),
    mean_tmean = mean(apparent_tmean),
    .groups = "drop"
  ) %>%
  filter(duration >= 3) # enforce 3+ days
```

###97.5th mininum heatwave days + events 
```{r}
#tally heatwave days at 85th percentile mean
hw_975_min <- daymet_heatwaves %>%
  filter(month >= 4 & month < 10) %>% #restrict to SF summer dates to avoid seasonal bias
  mutate(hw = apparent_tmin > p97.5_min) 

hw_975_min$zip_code <- as.character(hw_975_min$zip_code)

hw_975_min <- hw_975_min %>%
  group_by(site, zip_code) %>%
  mutate(event_id = rleid(hw))

#summarise TRUE runs and 2+ day heatwaves
hwe_2day_975_min <- hw_975_min %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(),
    start_date = min(date),
    end_date = max(date),
    mean_tmin = mean(apparent_tmin),
    .groups = "drop"
  ) %>%
  filter(duration >= 2) # enforce 2+ days

#summarise TRUE runs and 3+ day heatwaves
hwe_3day_975_min <- hw_975_min %>%
  filter(hw) %>% #keeps only heatwave days
  group_by(site, event_id, zip_code, year) %>%
  summarise(
    duration = n(), #adjust to just "duration" if 
    start_date = min(date),
    end_date = max(date),
    mean_tmin = mean(apparent_tmin),
    .groups = "drop"
  ) %>%
  filter(duration >= 3) # enforce 3+ days
```

##cleaning 911 calls 
```{r}
per_day <- medical_incidents |>
  mutate(Call.Date = mdy(Call.Date)) |>
  count(Zipcode.of.Incident, Call.Date) 

per_day <- per_day |>
  mutate(year = year(Call.Date),
         month = month(Call.Date),
         Zipcode.of.Incident = as.character(Zipcode.of.Incident)) |>
  relocate(year, .after = "Call.Date") |> 
  filter(month >= 4 & month < 10) %>% 
  drop_na(Zipcode.of.Incident)
```

#merge heatwaves and 911 calls 
```{r}

#85 mean 
overall_current_85_mean <- left_join(
  per_day,
  hw_85_mean,
  by = c(
    "Call.Date" = "date",
    "Zipcode.of.Incident" = "zip_code",
    "year" = "year",
    "month" = "month"
  ),
  na_matches = "never"
)

overall_current_85_mean <- overall_current_85_mean %>% 
  mutate(day_abbr = format(as.Date(Call.Date), "%a"),
         date = Call.Date)

#85 min
overall_current_85_min <- left_join(
  per_day,
  hw_85_min,
  by = c(
    "Call.Date" = "date",
    "Zipcode.of.Incident" = "zip_code",
    "year" = "year",
    "month" = "month"
  ),
  na_matches = "never"
)

overall_current_85_min <- overall_current_85_min %>% 
  mutate(day_abbr = format(as.Date(Call.Date), "%a"),
         date = Call.Date)

#97.5th mean 
overall_current_975_mean <- left_join(
  per_day,
  hw_975_mean,
  by = c(
    "Call.Date" = "date",
    "Zipcode.of.Incident" = "zip_code",
    "year" = "year",
    "month" = "month"
  ),
  na_matches = "never"
)

overall_current_975_mean <- overall_current_975_mean %>% 
  mutate(day_abbr = format(as.Date(Call.Date), "%a"),
         date = Call.Date)

#97.5th min 
overall_current_975_min <- left_join(
  per_day,
  hw_975_min,
  by = c(
    "Call.Date" = "date",
    "Zipcode.of.Incident" = "zip_code",
    "year" = "year",
    "month" = "month"
  ),
  na_matches = "never"
)

overall_current_975_min <- overall_current_975_min %>% 
  mutate(day_abbr = format(as.Date(Call.Date), "%a"),
         date = Call.Date)
```

#85th percentiles 

##85th mean 
```{r}
dat_85_mean <- overall_current_85_mean %>%
  mutate(
    Call.Date = as.Date(Call.Date),
    year = as.integer(format(Call.Date, "%Y")),
    month = as.integer(format(Call.Date, "%m")),
    dow = format(Call.Date, "%a"),
    hw = as.integer(hw),
    cc_strata = interaction(Zipcode.of.Incident, year, month, dow, drop = TRUE)
  ) %>%
  group_by(Zipcode.of.Incident) %>%
  arrange(Call.Date, .by_group = TRUE) %>%
  mutate(
    hw_lag1 = lag(hw, 1),
    hw_lag2 = lag(hw, 2),
    hw_lag3 = lag(hw, 3)
  ) %>%
  ungroup()

fit_lag_85_mean <- gnm(
  n ~ hw + hw_lag1 + hw_lag2 + hw_lag3,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat_85_mean
)

summary(fit_lag_85_mean)

beta <- coef(fit_lag_85_mean)[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]
se <- sqrt(diag(vcov(fit_lag_85_mean)))[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]

results_85_mean <- data.frame(
  term = c("85 Heat wave (lag 0)", "85 Heat wave (lag 1)", "85 Heat wave (lag 2)", "85 Heat wave (lag 3)"),
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

results_85_mean
```

##85th mininmum 
```{r}
dat_85_min <- overall_current_85_min %>%
  mutate(
    Call.Date = as.Date(Call.Date),
    year = as.integer(format(Call.Date, "%Y")),
    month = as.integer(format(Call.Date, "%m")),
    dow = format(Call.Date, "%a"),
    hw = as.integer(hw),
    cc_strata = interaction(Zipcode.of.Incident, year, month, dow, drop = TRUE)
  ) %>%
  group_by(Zipcode.of.Incident) %>%
  arrange(Call.Date, .by_group = TRUE) %>%
  mutate(
    hw_lag1 = lag(hw, 1),
    hw_lag2 = lag(hw, 2),
    hw_lag3 = lag(hw, 3)
  ) %>%
  ungroup()

fit_lag_85_min <- gnm(
  n ~ hw + hw_lag1 + hw_lag2 + hw_lag3,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat_85_min
)

summary(fit_lag_85_min)

beta <- coef(fit_lag_85_min)[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]
se <- sqrt(diag(vcov(fit_lag_85_min)))[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]

results_85_min <- data.frame(
  term = c("85 Heat wave (lag 0)", "85 Heat wave (lag 1)", "85 Heat wave (lag 2)", "85 Heat wave (lag 3)"),
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

results_85_min
```

#97.5th percentiles 

##97.5th mean 
```{r}
dat_975_mean <- overall_current_975_mean %>%
  mutate(
    Call.Date = as.Date(Call.Date),
    year = as.integer(format(Call.Date, "%Y")),
    month = as.integer(format(Call.Date, "%m")),
    dow = format(Call.Date, "%a"),
    hw = as.integer(hw),
    cc_strata = interaction(Zipcode.of.Incident, year, month, dow, drop = TRUE)
  ) %>%
  group_by(Zipcode.of.Incident) %>%
  arrange(Call.Date, .by_group = TRUE) %>%
  mutate(
    hw_lag1 = lag(hw, 1),
    hw_lag2 = lag(hw, 2),
    hw_lag3 = lag(hw, 3)
  ) %>%
  ungroup()

fit_lag_975_mean <- gnm(
  n ~ hw + hw_lag1 + hw_lag2 + hw_lag3,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat_975_mean
)

summary(fit_lag_975_mean)

beta <- coef(fit_lag_975_mean)[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]
se <- sqrt(diag(vcov(fit_lag_975_mean)))[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]

results_975_mean <- data.frame(
  term = c("97.5 Heat wave (lag 0)", "97.5 Heat wave (lag 1)", "97.5 Heat wave (lag 2)", "97.5 Heat wave (lag 3)"),
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

results_975_mean
```

##97.5th minimum 
```{r}
dat_975_min <- overall_current_975_mean %>%
  mutate(
    Call.Date = as.Date(Call.Date),
    year = as.integer(format(Call.Date, "%Y")),
    month = as.integer(format(Call.Date, "%m")),
    dow = format(Call.Date, "%a"),
    hw = as.integer(hw),
    cc_strata = interaction(Zipcode.of.Incident, year, month, dow, drop = TRUE)
  ) %>%
  group_by(Zipcode.of.Incident) %>%
  arrange(Call.Date, .by_group = TRUE) %>%
  mutate(
    hw_lag1 = lag(hw, 1),
    hw_lag2 = lag(hw, 2),
    hw_lag3 = lag(hw, 3)
  ) %>%
  ungroup()

fit_lag_975_min <- gnm(
  n ~ hw + hw_lag1 + hw_lag2 + hw_lag3,
  eliminate = cc_strata,
  family = poisson(link = "log"),
  data = dat_975_min
)

summary(fit_lag_975_min)

beta <- coef(fit_lag_975_min)[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]
se <- sqrt(diag(vcov(fit_lag_975_min)))[c("hw", "hw_lag1", "hw_lag2", "hw_lag3")]

results_975_min <- data.frame(
  term = c("97.5 Heat wave (lag 0)", "97.5 Heat wave (lag 1)", "97.5 Heat wave (lag 2)", "97.5 Heat wave (lag 3)"),
  IRR = exp(beta),
  LCL = exp(beta - 1.96 * se),
  UCL = exp(beta + 1.96 * se)
)

results_975_mean
```
#IRR actual percentages 
```{r}
all_results_mean <- rbind(results_85_mean, results_975_mean)

all_results_mean <- all_results_mean %>% 
  mutate(
    IRR_percent = (IRR - 1)*100
  )

all_results_min <- rbind(results_85_min, results_975_min)

all_results_min <- all_results_min %>% 
  mutate(
    IRR_percent = (IRR - 1)*100
  )

```

